---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Reading data from csv

```{r echo=TRUE}
df <- read.csv("activity.csv")
head(df)
print(colnames(df))
options(scipen = 100)
```



## What is mean total number of steps taken per day?



```{r echo=TRUE}
total.steps <- tapply(df$steps, df$date, sum, na.rm = T)
hist(total.steps, 
	main = "Histogram of total steps per day", 
	xlab = "Total steps per day", 
	ylab = "Frequency", 
	breaks = 20,
	col = "green"
)
steps.mean <- mean(total.steps, na.rm = T)
steps.median <- median(total.steps, na.rm = T)
```

The mean number of steps taken each day is `r round(steps.mean, 0)`, the median number is `r round(steps.median, 0)`. 


## What is the average daily activity pattern?

```{r echo=TRUE}
steps.interval <- aggregate(df$steps, list(df$interval), mean, na.rm = T)
colnames(steps.interval) <- c("interval", "steps")
plot(steps.interval, type = "l",
	main = "Average steps by 5-minute interval",
	xlab = "Interval",
	ylab = "Average number of steps",
)
```

5-minute interval, on average across all the days in the dataset that contains the maximum number of steps

```{r}
max <- steps.interval[which.max(steps.interval$steps),]
```

Maximum average is in the interval `r max$interval`

## Imputing missing values

Number of NA Values

```{r}
nrow(df[is.na(df$steps), ])
```

Loading important libraries. Imputing data using impute function. 

```{r}
library(scales)
library(Hmisc)

imputedData <- df
imputedData$steps <- impute(df$steps, fun = mean)
head(imputedData)
```

Make a histogram of the total number of steps taken each day 

```{r}
total.na.steps <- tapply(imputedData$steps, imputedData$date, sum)
hist(total.na.steps, 
	main = "Histogram of total steps per day", 
	xlab = "Total steps per day", 
	ylab = "Frequency", 
	breaks = 20,
	col = "red"
)
steps.mean <- mean(total.na.steps)
steps.median <- median(total.na.steps)
```

The mean number of steps taken each day is `r round(steps.mean, 0)`, the median number is `r round(steps.median, 0)`. 


## Are there differences in activity patterns between weekdays and weekends?

```{r}
imputedData$datetype <- ifelse(as.POSIXlt(imputedData$date)$wday %in% c(0,6), 'weekend', 'weekday') 
```

### Make a panel plot

```{r}
averageActivity <- aggregate(steps ~ interval + datetype, data = imputedData, mean)
ggplot(averageActivity, aes(interval, steps)) + 
  geom_line() + 
  facet_grid(datetype ~ .) + 
  xlab("5-minute interval") + 
  ylab("avarage number of steps")
```

